---
title: "Data Science with R (Day 4)"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    code_folding: hide
    fig_width: 8
    fig_height: 6
    fig_caption: true
    css: styles.css
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
```

# Learning Objectives

- Repetition on Data Manipulation and Plotting

- Document Styles with YAML

- Web Scraping

- Functions

- Formatting of Rmarkdown

- CSS

- Integrate math formulas

- Tabbed sections

# Open Questions

# Last exercise

discuss last exercise

# Repetition on Data Manipulation and Plotting

The dataset includes data on affairs of politicians. We will try to find out, if and how the number of affairs is related to sex, age, years of marriage, having children, being religious, educational level, occupation.

If you are interested in the authors publication you can read it here

[https://fairmodel.econ.yale.edu/rayfair/pdf/1978a200.pdf]

## Data Preparation

### Data Import 

We load data from this location.

[http://koaning.io/theme/data/affairs.csv]

```{r}
library(pacman)
p_load(tidyverse)
```

```{r}
url <- "http://koaning.io/theme/data/affairs.csv"
filename <- "Affairs.csv"
download.file(url = url, destfile = filename)
```

```{r}
affairs <- read_csv(file = filename, col_names = T)
```

### Data Understanding

We mMake ourselves familiar with data. Which columns are included? What is their range, min, max, median?

```{r}
summary(affairs)
```


## Analysis

### Hypothesis 1: Men are more likely to have affairs than women.

Variable "nbaffairs" defines the intensity of the affair in terms of having intercourse (0 = none, 12 = daily). Please create an additional (binary) variable "affair_yes_no", that is 1 for "nbaffairs" > 0 and 0 for "nbaffairs" == 0.

Please plot the result in a suitable graph.

```{r}
aff_likelihood <- affairs %>% 
  mutate (affair_yes_no = ifelse(nbaffairs > 0, 1, 0)) %>% 
  group_by(sex) %>% 
  summarise (affair = sum (affair_yes_no),
             count = n(),
             affair_ratio = affair / count * 100)

g <- ggplot(aff_likelihood, aes(sex, affair_ratio))
g <- g + geom_col(fill = "blue")
g <- g + labs (title = "Likelihood of Having an Affair", x = "Sex", y = "Proportion [%]")
g <- g + theme_bw()
g
```

### Hypothesis 2: Religious Politicians are less likely to have an affair.

```{r}
affairs <- affairs %>% 
  mutate (affair_yes_no = ifelse(nbaffairs > 0, 1, 0)) 

aff_likelihood <- affairs %>%
  group_by(sex, religious) %>% 
  summarise (affair = sum (affair_yes_no),
             count = n(),
             affair_ratio = affair / count * 100)

g <- ggplot(aff_likelihood, aes(sex, affair_ratio, fill = factor(religious)))
g <- g + geom_col(position = "dodge")
g <- g + labs (title = "Likelihood of Having an Affair for Sex and Religiousness", x = "Sex", y = "Proportion [%]")
g <- g + scale_fill_discrete(name = "Religiousness")
g <- g + theme_bw()
g
```

### Hypothesis 3a: Older, Male Politicians are more likely to have an affair.

Age should not be used directly, but rather binned for years 21-30, 31-40, ...

A small number of politicians with affairs is not representative, so please filter the dataset, if there are less than 3.

```{r}
aff_age <- affairs %>% 
  mutate (age_bin = paste0(age %/% 10 * 10 + 1, "-", age %/% 10 * 10 + 10)) %>% 
  group_by(age_bin, sex) %>% 
  summarise(affair_count = sum (affair_yes_no),
            count = n(),
            affair_prob = affair_count / count * 100) %>% 
  filter (affair_count > 3)

g <- ggplot(aff_age, aes(x = age_bin, y = affair_prob, fill = sex))
g <- g + geom_col(position = "dodge")
g <- g + labs (title = "Impact of Age and Sex on Having an Affair", x = "Age Class", y = "Affair-Probability [%]")
g <- g + theme_bw()
g
```

### Hypothesis 3b: identical to 3a

Here, please use a heatmap for visualisation.

```{r}
library(pheatmap)
aff_age_spread <- aff_age %>% 
  select (-affair_count, -count) %>% 
  spread (key = "age_bin", value = "affair_prob")
annotation <- aff_age_spread$sex

pheatmap(aff_age_spread[, 2:5], 
         cluster_rows = F, 
         cluster_cols = F, 
         main = "Affairs for Age and Sex", 
         display_numbers = T, 
         number_format = "%.0f")
```

### Hypothesis 4: Ratio of Affairs increases with Years of Marriage.

```{r}
aff_ym_nbaffairs <- affairs %>% 
  group_by(ym) %>% 
  summarise(affairs = sum(affair_yes_no),
            count = n(),
            affair_ratio = affairs / count * 100
            )

g <- ggplot(aff_ym_nbaffairs, aes(ym, affairs))
g <- g + geom_smooth(method = "lm", se = F)
g <- g + geom_point()
g <- g + labs (title = "Ratio of Having an Affair and Years of Marriage", 
               x = "Years of Marriage [-]", 
               y = "Ratio of Having an Affair [%]")
g <- g + theme_bw()
g <- g + geom_text(aes(label = paste(round(affair_ratio, 1), "%")), vjust = 1.1)
g <- g + scale_x_continuous(breaks = seq(0, 16, 2))
g
```

# Web Scraping

```{r}
library(rvest)
```

```{r}
url <- "https://en.wikipedia.org/wiki/Wind_power_by_country"
xpath <- '//*[@id="mw-content-text"]/div/table[3]'

table <- html(url) %>% 
    html_node(xpath =  xpath) %>% 
    html_table(fill = T)

```

# Functions

We will define a function for calculating AEP for a given power curve.

```{r}
Aep <- function(ws_mean = 8 , k = 2, ws, power, cutout, step_size) {
  N_h <- 8760  # hours per year
  df <- data.frame(ws = ws, 
                   power = power)
  df$F <- 1 - exp(-pi/4*(df$ws/ws_mean)**2)
  N <- nrow(df)  # number of bins
  df$aep <- NA
  df$aep[1] <- df$F[1] * df$power[1] / 2 * N_h
  for (i in 2:N) {
    df$aep[i] <- N_h * (df$F[i]-df$F[i-1])*(df$power[i-1]+df$power[i])/2 
  }
  aep_sum <- sum(df$aep, na.rm = T)
  return(aep_sum)
}
```

# Document Styles with YAML

YAML stands for Yet Another Meta Language. It defines the look and feel of HTML-document.

More information can be found [http://rmarkdown.rstudio.com/html_document_format.html].

# Formatting of Rmarkdown Text

Rmarkdown text can be formatted. All formats are described in Rmarkdown Reference [https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf].


# Custom Style Sheets

You can manually adapt Rmarkdown layout with CSS.

More information to be found at:
<https://rmarkdown.rstudio.com/html_document_format.html>

# Math formulas

You can add mathematical formulas as with LaTex.

$$ws_{norm} = ws_{meas} \cdot \frac{\rho_{meas}}{\rho_{standard}}^{0.333}$$

# Tabbed sections {.tabset}

## Tab1 

some content here.

## Tab2

some more content here.

## Tab3

much more content here.